---
title: "OpenStreetMap"
---

# Extracting data from OpenStreetMap

## Installing OpenStreetMap packages

Filtering and extracting contents from OpenStreetMap dump files is facilitated by dedicated utilities. In this project, [Osmconvert](http://wiki.openstreetmap.org/wiki/Osmconvert) and [Osmfilter](http://wiki.openstreetmap.org/wiki/Osmfilter) have been used to filter relevant data.

```{r eval=FALSE, include=FALSE}
# osmosis not needed? (Osmosis)[http://wiki.openstreetmap.org/wiki/Osmosis/Installation] 
# following code works on linux, for other OS see:
#  http://wiki.openstreetmap.org/wiki/Osmosis/Installation
dir.create(path = "osmosis", showWarnings = FALSE)

if (file.exists(file.path("osmosis", "osmosis-latest.tgz"))==FALSE) {
  download.file(url = "http://bretth.dev.openstreetmap.org/osmosis-build/osmosis-latest.tgz",
                destfile = file.path("osmosis", "osmosis-latest.tgz"))
  untar(tarfile = file.path("osmosis", "osmosis-latest.tgz"), exdir = "osmosis")
  system(command = "chmod a+x osmosis/bin/osmosis")
}
```

```{r}
# install osmfilter
if (file.exists("osmfilter")==FALSE) {
  system(command = "wget -O - http://m.m.i24.cc/osmfilter.c |cc -x c - -O3 -o osmfilter")
}

# install osmconvert
if (file.exists(file.path("osmconvert64"))==FALSE) {
  download.file(url = "http://m.m.i24.cc/osmconvert64", destfile = file.path("osmconvert64"))
  system(command = "chmod +x osmconvert64")
}


```

## Download OpenStreetMap data

It is possible to download all available data for selected countries from [online repositories](http://download.geofabrik.de/europe.html).

```{r}
dir.create(path = "data", showWarnings = FALSE)
dir.create(path = file.path("data", "pbf"), showWarnings = FALSE)


for (i in countries) {
  if (file.exists(file.path("data", "pbf", paste0(i, "-latest.osm.pbf")))==FALSE) {
    download.file(url = paste0("http://download.geofabrik.de/europe/", i, "-latest.osm.pbf"),
                  destfile = file.path("data", "pbf", paste0(i, "-latest.osm.pbf")))
  }
}

```

## Keep only street names

Through the above-mentioned utilities, all names of street found in the region have been extracted from the dump and converted to .cvs format. This has been accomplished by filtering only the names of objects recorded under the key `highway` in OpenStreetMap, which according to [OpenStreetMap's wiki](http://wiki.openstreetmap.org/wiki/Key:highway) and in spite of its name, "is the main key used for identifying any kind of road, street or path."

At this stage, only streets have been included in the filter, but other OpenStreetMap categories may include relevant data, e.g. [`Tag:place=square`](http://wiki.openstreetmap.org/wiki/Tag:place%3Dsquare), or [`Tag:leisure=park`](http://wiki.openstreetmap.org/wiki/Tag:leisure%3Dpark).

```{r}
# convert to a format that can be read by osmfilter, and remove author data to reduce file size
dir.create(path = file.path("data", "o5m"), showWarnings = FALSE)

for (i in countries) {
  if (file.exists(file.path("data", "o5m", paste0(i, "-latest.o5m")))==FALSE) {
    system(paste0('./osmconvert64 data/pbf/', i, 
                  '-latest.osm.pbf --drop-version --out-o5m -o=data/o5m/', i,
                  '-latest.o5m'))
  }
}

# filter only streets
dir.create(path = file.path("data", "o5m-streets"), showWarnings = FALSE)

for (i in countries) {
  if (file.exists(file.path("data", "o5m-streets", paste0(i, "-streets.o5m")))==FALSE) {
    system(paste0('./osmfilter data/o5m/', i, '-latest.o5m --keep="highway=*" --drop-version > ', 'data/o5m-streets/', i, '-streets.o5m'))
  }
}

# export to csv only street type, name, and lon/lat

dir.create(path = file.path("data", "csv-streets"), showWarnings = FALSE)

for (i in countries) {
  if (file.exists(file.path("data", "csv-streets", paste0(i, "-streets.csv")))==FALSE) {
    system(paste0('./osmconvert64 data/o5m-streets/', i,
                  '-streets.o5m --all-to-nodes --csv="@id @lat @lon highway name" > data/csv-streets/', i,
                  '-streets.csv', " --csv-separator='; '"))
  }
}

```

## Keep only street names including reference to Tito

Once all street names have been extracted, only street names inlcuding reference to Tito have been kept for further analysis.

Criteria for filters:

- ends with 'Tito'
- ends with 'Tita'
- includes 'Titov'
- includes 'Тито' (and accordingly, also all 'Титов' etc., but excluding "Ктито*')

```{r message=FALSE, warning=FALSE}

StreetsSummary <- data_frame(Country = countries, TotalStreets = as.integer(NA), TitoStreets = as.integer(NA))

FindTito <- "Tito$|Tita$|Titov|Титов|(?<!К)Тито"
tito_all <- data_frame()

for (i in countries) {
  # Import from csv
  temp_streets <- read_delim(file = file.path("data", "csv-streets", paste0(i, "-streets.csv")),
                             delim = "; ",
                             col_names = FALSE, 
                             locale = locale(decimal_mark = "."),
                             trim_ws = TRUE)
  # Store total number of streets for each country
  StreetsSummary$TotalStreets[StreetsSummary$Country==i] <- nrow(temp_streets)
  # Filter only Tito's streets
  temp_tito <- temp_streets %>% 
    filter(is.na(X5)==FALSE) %>% 
    filter(stringr::str_detect(string = X5, pattern = stringr::regex(pattern = FindTito, ignore_case = TRUE))) %>%
    transmute(lat = X2, lon = X3, streetname = X5)  %>%
    # remove duplicate location with which have similar coordinates (same first decimal)
    mutate(lonShort = format(lon, digit = 3), latShort = format(lat, digit = 3)) %>%
    distinct(lonShort, latShort, .keep_all = TRUE) %>% 
    select(-lonShort, -latShort) %>% 
    mutate(country = i)
  # Store number of Tito's streets in given country
  StreetsSummary$TitoStreets[StreetsSummary$Country==i] <- nrow(temp_tito)
  # Merge table for each country
  tito_all <- bind_rows(tito_all, temp_tito)
}

tito_all <- tito_all %>% distinct()

# tito_all <- tito_all %>% 
#   mutate(lonShort = format(lon, digit = 4), latShort = format(lat, digit = 4)) %>%
#   distinct(lonShort, latShort, .keep_all = TRUE) %>% 
#   select(-lonShort, -latShort)

ExportData(data = tito_all, filename = "OSM_tito_all")
```

## Preliminary results based on OpenStreetMap

**As appears from the following sections, these data include substantial inaccuracies. They are reported here as preliminary results based on OpenStreetMap. See the following sections for more accurate data.**

### Number of streets dedicated to Tito

```{r}
StreetsSummary %>%
  arrange(TitoStreets) %>% 
  mutate(Country = toupper(Country)) %>% 
  mutate(Country = forcats::fct_inorder(Country)) %>% 
  ggplot(aes(x = Country, y = TitoStreets, label = TitoStreets)) +
  geom_col() +
  scale_y_continuous(name = "", limits = c(0, 30)) +
  scale_x_discrete(name = "") +
  geom_text(hjust = -0.1) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Number of streets dedicated to Tito",
       subtitle = "Based on OpenStreetMaps data")

```

```{r}
ShowTable(data = StreetsSummary %>% arrange(desc(TitoStreets)))
```

## Putting streets dedicated to Tito (according to OpenStreetMap) on a map

```{r message=FALSE, warning=FALSE}
# Setting the coordinates
wb <- c(left = 12, bottom = 40, right = 26, top = 48)

# Preparing the canvas, and forcing cache
if (file.exists(file.path("temp", "mapTonerLite.rds"))==FALSE) {
  saveRDS(object = ggmap::get_stamenmap(bbox = wb, zoom = 6, maptype = "toner-lite") %>% ggmap(), file = file.path("temp", "mapTonerLite.rds"))
}

mapTonerLite <- readRDS(file = file.path("temp", "mapTonerLite.rds"))

OSM_TitoTonerLight <-  mapTonerLite + geom_point(data=tito_all, aes(x=lon, y=lat), color="red", size=2, alpha=0.5) +
  labs(x = '', y = '') + 
  labs(title = "Streets and squares dedicated to Tito", 
       subtitle = paste("based on OpenStreetMap"))

OSM_TitoTonerLight

ExportGraph(graph = OSM_TitoTonerLight, filename = "OSM_TitoTonerLight")  
```


```{r}
# Preparing the canvas  
dir.create(path = "temp", showWarnings = FALSE)

## forcing cache to prevent reloading map from Google Maps each time
if (file.exists(file.path("temp", "mapG.rds"))==FALSE) {
  saveRDS(object = get_googlemap("Sarajevo", scale = 2, zoom = 6), file = file.path("temp", "mapG.rds"))
}

mapG <- readRDS(file = file.path("temp", "mapG.rds"))

ggmap(mapG) + 
  geom_point(data=tito_all, aes(x=lon, y=lat), color="red", size=2, alpha=0.5) +
  labs(x = '', y = '') +
  labs(title = "Streets and squares dedicated to Tito", 
       subtitle = paste("Based on OpenStreetMap data"))

ExportGraph(filename = "OSM_TitoGmaps")  

# # Preparing the canvas
# mapWatercolor <- get_stamenmap(bbox = wb, zoom = 7, maptype = "watercolor") %>% ggmap()
# # Adding the dots
# mapWatercolor + geom_point(data=tito_all, aes(x=lon, y=lat), color="red", size=2, alpha=0.5) +
#   labs(x = '', y = '') + ggtitle("Streets dedicated to Tito")
# ggsave("TitoWatercolor.png")

```


## Alternative density maps

Testing alternative visualisations.

```{r}
mapTonerLite + 
  stat_density2d(
    aes(x = lon, y = lat, fill = ..level..,  alpha = ..level..),
    size = 2, bins = 4, data = tito_all,
    geom = "polygon"
)
```


## Summary of all relevant street names found on OpenStreetMap

The following table presents all versions of street names including reference to Tito found in each of the countries included in the analysis.

```{r}
tito_all_streetname <- tito_all %>% select(country, streetname) %>% group_by(country, streetname) %>% count() %>% arrange(country, desc(n), streetname)

ShowTable(data = tito_all_streetname)

ExportData(data = tito_all_streetname, filename = "tito_all_streetname")

```
