---
title: "Analysis"
---

# Where is Tito?


## See the results on a map


```{r}
TitoGmapsResults <- ImportData(filename = "TitoGmapsResults")
# Preparing the canvas, and forcing cache
wb <- c(left = 12, bottom = 40, right = 26, top = 48)
if (file.exists(file.path("temp", "mapTonerLite.rds"))==FALSE) {
  saveRDS(object = ggmap::get_stamenmap(bbox = wb, zoom = 6, maptype = "toner-lite") %>% ggmap(), file = file.path("temp", "mapTonerLite.rds"))
}

mapTonerLite <- readRDS(file = file.path("temp", "mapTonerLite.rds"))

mapTonerLite + geom_point(data=TitoGmapsResults, aes(x=lon, y=lat), color="brown", size=2, alpha=0.5) +
  labs(x = '', y = '') + labs(title = "Streets and squares dedicated to Tito", subtitle = paste("(based on Google Maps data as of", Sys.Date(), ")"))



#readRDS(file = file.path("graphs", "OSM_TitoTonerLight.rds"))
```

Comparing with OpenStreetMaps results

```{r}
mapTonerLite + geom_point(data=ImportData("OSM_tito_all"), aes(x=lon, y=lat), color="orange", size=2, alpha=0.5) +
  labs(x = '', y = '') + labs(title = "Streets and squares dedicated to Tito", subtitle = paste("(based on OpenStreetMap data as of", Sys.Date(), ")"))

```

On the same map:

```{r}
OSM_gmaps_Tito <- bind_rows(TitoGmapsResults, ImportData("OSM_tito_all") %>% select(lon, lat), .id = "Source")

mapTonerLite + geom_point(data=OSM_gmaps_Tito, aes(x=lon, y=lat, color = Source), size=2, alpha=0.5) +
  labs(x = '', y = '') + labs(title = "Streets and squares dedicated to Tito")


```


## Density visualisations

```{r}
mapTonerLite + 
  stat_density2d(
    aes(x = lon, y = lat, fill = ..level..,  alpha = ..level..),
    size = 2, bins = 4, data = TitoGmapsResults,
    geom = "polygon"
)

mapTonerLite + 
  stat_density2d(
    aes(x = lon, y = lat, fill = ..level..,  alpha = ..level..),
    size = 2, bins = 6, data = TitoGmapsResults,
    geom = "polygon"
)

mapTonerLite + 
  stat_density2d(
    aes(x = lon, y = lat, fill = ..level..,  alpha = ..level..),
    size = 2, bins = 8, data = TitoGmapsResults,
    geom = "polygon"
)


mapTonerLite + stat_bin2d(
  aes(x = lon, y = lat),
  size = .5, bins = 20, alpha = 1/2,
  data = TitoGmapsResults) +
  scale_fill_gradient(low = "#ffffb2", high = "#bd0026")

```

## By country / Administrative level

At the sub-country level, Google Maps records two levels of administrative sub-divisions. Unfortunately, not all streets are recorded with the respective sub-division, and the categorisation does not seem to be consistent.

```{r}

TitoGmapsMore <- bind_rows(ImportData(filename = "titovResults"), ImportData("marsalaTitaResults")) %>% 
  filter(type=="route", country != "Italy") %>% # exclude non-YU and non streets/squares
  filter(grepl(pattern = "Tit|Tит", x = route)) %>% # remove most non-tito
  filter(!grepl(pattern = "Strozzi|Brezova", x = route)) %>%  # remove remaining non-Tito
  distinct(address, .keep_all = TRUE) %>% # remove those with same address
  distinct(locality, route, .keep_all = TRUE) %>% #remove same locality, same street name
  distinct(lon, lat, route, administrative_area_level_1, administrative_area_level_2, country)

ShowTable(data = TitoGmapsMore %>% count(country, sort = TRUE), caption = "Streets and squares dedicated to Tito in the former Yugoslavia, by country, according to Google Maps/#FindingTito")

ExportData(data = TitoGmapsMore %>% count(country, sort = TRUE), filename = "TitoGmapsByCountry")

```

```{r}
TitoGmapsByCountry_gg <- ImportData(filename = "TitoGmapsByCountry") %>% mutate(country = forcats::fct_inorder(country)) %>% 
  ggplot(mapping = aes(x = country, y = n, label = n)) +
  geom_col() +
  geom_text(nudge_y = 5) +
  scale_x_discrete(name = "") +
  scale_y_continuous(name = "") +
  theme_minimal() +
  coord_flip() +
  labs(title = "Number of streets dedicated to Tito in the former Yugoslavia (by country)", subtitle = "According to Goole Maps/FindingTito")

TitoGmapsByCountry_gg
ExportGraph(graph = TitoGmapsByCountry_gg, filename = "TitoGmapsByCountry_gg")
```


```{r}
ShowTable(TitoGmapsMore %>% count(administrative_area_level_1, country, sort = TRUE))
```

```{r}
ShowTable(TitoGmapsMore %>% count(administrative_area_level_2, country, sort = TRUE))
```



## By street name

This is not very useful, however, since it mostly points at the different ways in which Google Maps stores street names.

```{r}
ShowTable(TitoGmapsResults %>% count(route, sort = TRUE))

```

```{r}
mapTonerLite + 
  geom_point(data=TitoGmapsResults, aes(x=lon, y=lat, color = route), size=2, alpha=0.5) +
  labs(x = '', y = '') + labs(title = "Streets and squares dedicated to Tito")


mapTonerLite + stat_bin2d(
  aes(x = lon, y = lat, fill = route, colour = route),
  size = .5, bins = 20, alpha = 1/2,
  data = TitoGmapsResults)
```

Keeping only street names found more than once.

```{r}
mapTonerLite + stat_bin2d(
  aes(x = lon, y = lat, fill = route, colour = route),
  size = .5, bins = 20, alpha = 1/2,
  data = left_join(TitoGmapsResults, TitoGmapsResults %>% count(route)) %>% filter(n>1))
```
Or more than twice.

```{r}
mapTonerLite + stat_bin2d(
  aes(x = lon, y = lat, fill = route, colour = route),
  size = .5, bins = 20, alpha = 1/2,
  data = left_join(TitoGmapsResults, TitoGmapsResults %>% count(route)) %>% filter(n>2))

mapTonerLite + 
  geom_point(data=left_join(TitoGmapsResults, TitoGmapsResults %>% count(route)) %>% filter(n>2), aes(x=lon, y=lat, color = route), size=2, alpha=0.5) +
  labs(x = '', y = '') + labs(title = "Streets and squares dedicated to Tito")

```

